Drought indices of water demand and supply and their impact on LULCC in Chile from ERA5-Land and MODIS

### 2024-01-26 article/paper_drought_Chile2023.qmd
actualizar repo

In contrast to our observations, @Venegas2022, who investigated Cryptocarya alba and Beilschmiedia miersii (both members of the Lauraceae family) inhabiting sclerophyllous forests in Chile, concluded that the overall development of the trees had decelerated, potentially indicating alterations in the inherent dynamics of their respective forests. They attributed it to the cumulative effects of the unprecedented drought (i.e., hydrological drought). Nonetheless, the trees within sclerophyllous forests are highly heterogeneous, and the species analyzed by @Venegas2022 are not the predominant ones. Because of this, when we used MOD13A3 with a 1km spatial resolution to measure vegetation condition, it took the average condition of 1 square kilometer. Thus, we attribute that forest to being the most resistant to drought, due to the fact that most of the species comprising it are highly resilient to water scarcity. Then, for a study of the dynamics of forest to drought, e.g., sclerophyllous forest, on a local scale using remote sensing, we should use operational products at higher spatial resolutions as provided by the missions Landsat or Sentinel that could capture the heterogenity of the forest type.

## ERA5-Land derived soil moisture and vegetation productivity

@@

>>>>

In contrast to our observations, @Venegas2022, who investigated Cryptocarya alba and Beilschmiedia miersii (both members of the Lauraceae family) inhabiting sclerophyllous forests in Chile, concluded that the overall development of the trees had decelerated, potentially indicating alterations in the inherent dynamics of their respective forests. They attributed it to the cumulative effects of the unprecedented drought (i.e., hydrological drought). Thus, we attribute that forest to being the most resistant to drought, due to the fact that most of the species comprising it are highly resilient to water scarcity. compared to the other landcover classes. Also, when we used MOD13A3 with a 1km spatial resolution to measure vegetation condition, it took the average condition of 1 square kilometer. Then, to study how a type of forest (sclerophyllous forest) changes in response to drought on a local level using remote sensing, we should use operational products with higher spatial resolutions, such as those from Landsat or Sentinel, which can show how different the forests are.

## Soil moisture, vegetation productivity, and agricultural drought.

The main external factors that affect biomass production by vegetation are ET and SM, and the rate of ET in turn depends on the availability of water storage in the root zone. Thus, soil moisture plays a key role in land carbon uptake and, consequently, in the production of biomass [@Humphrey2021]. Moreover, @Zhang2022 indicates there is a bidirectional causality between soil moisture and vegetation productivity. For this, lastly, some studies have redefined agricultural drought as soil moisture drought from the hydrological perspective [@Loon2016;@Samaniego2018]. Even though soil moisture is the external factor most determinant of vegetation biomass, there are multiple internal factors, such as species, physiological characteristics, and plant hydraulics, that would affect vegetation productivity. Because of that, we believe that agricultural drought, referring to the drought that impacts vegetation productivity, is the most proper term, as originally defined by @Wilhite1985.

Our results showed that the drought index derived from soil moisture (SSI) was the most effective in explaining vegetation productivity within landcover macroclasses. However, there is strong spatial variability throughout Chile and between classes, mainly attributable to climate heterogeneity, hydrological status, or vegetation resistance to water scarcity. The semi-arid "Norte Chico" and the Mediterranean "Centro" were where SSI had the best performance. Medium-term deficits of 12 months are more relevant in the response of vegetation, which decreases to the south, and in the case of croplands, they seem to react in a shorter time, with six months (SSI-6) in "Centro."

@Chatterjee2022

@@

## Early drought forecasting

### 2024-01-26 article/paper_drought_Chile2023.tex
actualizar repo

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, interior hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

\hypertarget{soil-moisture-vegetation-productivity-and-agricultural-drought.}{%
\subsection{Soil moisture, vegetation productivity, and agricultural
drought.}\label{soil-moisture-vegetation-productivity-and-agricultural-drought.}}

The main external factors that affect biomass production by vegetation
are ET and SM, and the rate of ET in turn depends on the availability of
water storage in the root zone. Thus, soil moisture plays a key role in
land carbon uptake and, consequently, in the production of biomass
\citep{Humphrey2021}. Moreover, \citep{Zhang2022} indicates there is a
bidirectional causality between soil moisture and vegetation
productivity. For this, lastly, are some studies that have redefined
agricultural drought as soil moisture drought from the hydrological
perspective \citep{Loon2016, Samaniego2018}. Even though soil moisture
is the external factor most determinant of vegetation biomass, there are
multiple internal factors, such as species, physiological
characteristics, and plant hydraulics, that would affect vegetation
productivity. Because of that, we believe that agricultural drought,
referring to the drought that impacts vegetation productivity, is the
most proper term, as originally defined by \citep{Wilhite1985}.

Our results showed that the drought index derived from soil moisture
(SSI) was the most effective in explaining vegetation productivity
within landcover macroclasses. However, there is strong spatial
variability throughout Chile, mainly attributable to the climate
heterogeneity.

\citep{Chatterjee2022}

@@

\hypertarget{early-drought-forecasting}{%
\subsection{Early drought forecasting}\label{early-drought-forecasting}}
